---
title: "Assignment 3"
author: "Esther Chen"
date: "2022-10-16"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(results = "hide", message = FALSE)
```

##Set Up
```{r}
#Load packages
library(rvest)
library(sf)
library(tidyverse)
library(jsonlite)
library(tidycensus)
library(ggplot2)
```

#Part 1

```{r}
#Read in data
emergencyVisits <- fromJSON("https://data.cityofnewyork.us/resource/2nwg-uqyg.json?$limit=3000")
```

##Question 1
```{r}
emergencyVisits$date <- as.Date(emergencyVisits$date)
```

```{r}
emergencyVisits$date
```

##Question 2
```{r}
head(emergencyVisits)
```
The number of ED visits is a character.

```{r}
#Change the character to integer
emergencyVisits$total_ed_visits <- strtoi(emergencyVisits$total_ed_visits)
```

##Question 3

```{r}
#Order by date
emergencyVisits <- arrange(emergencyVisits, date)
```
```{r}
#Most recent 300 visits
recentvisits <- emergencyVisits[2701:3000,]
```

##Question 4
```{r}
#Visits per zip code
totalrecentvisits <- recentvisits %>%
  group_by(mod_zcta) %>%
  summarise(Total = sum(total_ed_visits))

totalvisits <- emergencyVisits %>%
  group_by(mod_zcta) %>%
  summarise(Total = sum(total_ed_visits))
```

#Part 2
```{r}
#Map of zipcodes
NYCmap = st_read("C:/Users/jk/Documents/ZIP_CODE_040114.shp")

ggplot(sample_frac(NYCmap)) + geom_sf() 
```

#Part 3

##Question 6
```{r}
#Joining the data
NYCmapChr = mutate(NYCmap, mod_zcta = as.character(ZIPCODE))
NYCmapZip = select(NYCmapChr, mod_zcta, geometry)

recentvisitsmap <- merge(NYCmapZip, recentvisits, by="mod_zcta")

totalvisitsmap <-merge(NYCmapZip, totalvisits, by="mod_zcta")

totalrecentvisitsmap <- merge(NYCmapZip, totalrecentvisits)
```

##Question 7 
```{r}
plot(totalvisitsmap)
```

#Part 4

```{r}
#Set API key
census_api_key("98e447b6bc529e5f36793dc8ff9dac0b8981f37a")

#Get data from api
censusData = load_variables(2018, "acs5", cache=T)
populationData = get_acs(geography = "zcta",
variables = 'B01003_001',
geometry = FALSE)

#Remove the extra tag
populationData$mod_zcta = sub("ZCTA5 ", "", populationData$NAME)


#Rename the column
populationData =
populationData %>%
rename(population = estimate)
```

##Question 8
```{r}
#Merge the population data with my main dataset

populationvisits <- merge(totalvisitsmap, populationData, by="mod_zcta")

```

##Question 9
```{r}
#Calculate visits per inhabitant
populationvisits <- populationvisits %>%
  mutate(visits_per_inhabitant = Total/population)
```
#Part 5
```{r}
#Download poverty data from api
vars = c(poverty = 'B17001_002')
povertyData = get_acs(geography = "zcta",
variables = vars,
geometry = FALSE)

povertyData$mod_zcta = sub("ZCTA5 ", "", povertyData$NAME)

povertyData = povertyData %>% rename(povertyCount = estimate)
```
#Part 6
##Question 10
```{r}
#Inner join with population data
povertypopulation <- merge(populationvisits, povertyData, by="mod_zcta")
```

##Question 11
```{r}
#Calculate the poverty rate
povertypopulation = povertypopulation %>%
  mutate(poverty_rate = povertyCount/population)
```
#Part 7
##Question 12-14
```{r}
#Analyze the relationship between poverty rate and the number of visits

povertyVisits <- lm(visits_per_inhabitant ~ poverty_rate, data = povertypopulation)

summary(povertyVisits)

plot(x = povertypopulation$poverty_rate, y = povertypopulation$visits_per_inhabitant)
```

There is a slight positive relationship between the poverty rate of a zip code and the number of emergency room visits per person. 

#Part 8
#Question 15 and 16
```{r}
#Get data from API
vars = c(age = 'B01002_001')
ageData = get_acs(geography = "zcta", variables = vars, geometry = FALSE)

ageData$mod_zcta = sub("ZCTA5 ","",ageData$NAME)

ageData = ageData %>%
  rename(medianAge = estimate)

#Merge data
populationAge <- merge(populationvisits, ageData, by="mod_zcta")

#Analyze relationship
ageVisits <- lm(Total ~ medianAge, data = populationAge)

plot(populationAge$medianAge, populationAge$Total)

summary(ageVisits)
```
There is a negative association between the median age of a zip code and the total number of hospital visits in that zip code. 
#Part 9
##Question 17 and 18
```{r}
vars = c('B02001_002')
ethnicityData = get_acs(geography = "zcta", variables = vars, geometry = FALSE)
ethnicityData = ethnicityData %>% rename(whiteAloneCount = estimate)
ethnicityData$mod_zcta = sub("ZCTA5 ","",ethnicityData$NAME)

#Merge data and calculate percentage of population that stated they are "White Alone".

populationEthnicity <- merge(populationvisits, ethnicityData, by="mod_zcta") 

populationEthnicity <- populationEthnicity %>%
  mutate(percentWhite = whiteAloneCount/population)
```


```{r}
populationDat <- populationvisits %>%
  mutate(percentWhite = populationEthnicity$percentWhite)

populationDat <- populationDat %>%
  mutate(medianAge = populationAge$medianAge)

populationDat <- populationDat %>%
  mutate(povertyRate = povertypopulation$poverty_rate)
```
#Part 10
##Question 19 and 20
```{r}
#Multilinear regression and analysis
mlr <- lm(formula = visits_per_inhabitant ~ percentWhite + medianAge + povertyRate, data= populationDat)

summary(mlr)
```

By taking a look at the coefficients, we can see that being white and having a higher median age actually has slightly negative association with the rate of emergency room visits. And we can see that poverty has a positive relationship with emergency room visits. It looks like the factor with the largest association is poverty. I would say this makes sense because a person's ability to get preventative care is deeply related to their economic status. Therefore it could lead to more emergency and severe cases of illness before they finally seek treatment. 